---
title: "End of chapter 8&9"
author: "Zhijun Ma"
date: "February 5, 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load data and create dummy variable
```{r accidents}
accidents.df <- read.csv("AccidentsFull.csv")
View(accidents.df)
accidents.df$INJURY <- ifelse(accidents.df$MAX_SEV_IR == 0, "NO", "YES")
str(accidents.df)
```

## check the probability of a reported injury
```{r}
inj.tbl <- table(accidents.df$INJURY)
inj.tbl["YES"]
inj.prob<-prop.table(table(accidents.df$INJURY))
inj.prob
```

## as.factor
```{r}
for (i in c(1:dim(accidents.df)[2])){
  accidents.df[,i] <- as.factor(accidents.df[,i])
}
str(accidents.df)
```



## select first 12 rows and two predictors weather_r and traf_con_r
```{r pressure, echo=FALSE}
accidents.df1 <- accidents.df[c(1:12),c("WEATHER_R","TRAF_CON_R","INJURY")]
View(accidents.df1)

## pivot
ftable(accidents.df1)
```

## naive bayes %
```{r}

library(e1071)
library(caret)
library(klaR)
nb <- naiveBayes(INJURY ~ ., data = accidents.df1)
nb
predict(nb, newdata=accidents.df1, type="raw")       

nb2 <- train(INJURY ~ TRAF_CON_R + WEATHER_R, 
      data = accidents.df1, method = "nb")

predict(nb2, newdata = accidents.df1)
predict(nb2, newdata = accidents.df1,type = "raw")

                               
```
prop.table(table(accidents.df1$INJURY, accidents.df1$WEATHER_R), margin = 1)   

## exact bayes %
```{r}
#To find P(Injury=yes|WEATHER_R = 1, TRAF_CON_R =0):
#Numerator = (proportion of combination {WEATHER_R =1, TRAF_CON_R = 0}when Injury =  
#               yes) * (proportion of injuries in all cases)
#Denominator = proportion of combination {WEATHER_R =1, TRAF_CON_R = 0} in 
#all cases
#you want to reference the resulting table  for this based on the three variables
numerator1 <- 2/3 * 3/12
denominator1 <- 3/12
#the probability is given by ration
prob110 <- numerator1/denominator1
prob110
#so P(Injury=yes|WEATHER_R = 1, TRAF_CON_R =0) = 0.667
```

## classify the 12 accidents using these % and a cutoff of 0.5.
```{r}
prob.inj <- c(0.667,0.167,0,0,0.667,0.167,0.167,0.667,0.167,0.167,0.167,0)
accidents.df1$prob.inj <- prob.inj
accidents.df1
accidents.df1$pred.prob <- ifelse(accidents.df1$prob.inj > 0.5, "yes", "no")
accidents.df1
```

## use the entire dataset and partition 60-40
```{r}
library(e1071)
library(caret)
train.index <- sample(c(1:dim(accidents.df)[1]), dim(accidents.df)[1]*0.6)  
train.df <- accidents.df[train.index, ]
valid.df <- accidents.df[-train.index, ]
colnames(train.df)[1] <- "HOUR_I_R"
colnames(valid.df)[1] <- "HOUR_I_R"
vars <- c("INJURY","HOUR_I_R","ALIGN_I","WRK_ZONE","WKDY_I_R",
          "INT_HWY",	"LGTCON_I_R",	"PROFIL_I_R",	"SPD_LIM",	"SUR_COND",
          "TRAF_CON_R",	"TRAF_WAY",	"WEATHER_R")
nb <- naiveBayes(INJURY ~ ., data = train.df[, vars])
predt.df <- predict(nb, train.df[, vars])
confusionMatrix(predt.df,train.df$INJURY)
```

## overall error for validation set
```{r}
confusionMatrix(predict(nb, valid.df[, vars]),valid.df$INJURY)
```

## conditional probability output
```{r}
options(digits = 2)
nb

```

