---
title: '#4 Assignment - End of chapter 6&7'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Chapter 6.3
## exolore numerical predictor and response (fare), CORRELATION
```{r }
setwd("c:/Users/mazhi/Documents/R/560_data_mining")
airfares.df <- read.csv("Airfares.csv")
View(airfares.df)
cor(airfares.df[,c(5:6,9:13,16:18)])
```

## Scatterplots
```{r pressure, echo=FALSE}
library(ggplot2)
ggplot(airfares.df) + geom_point(aes(x = DISTANCE, y = FARE), colour = "navy", alpha = 0.7)
ggplot(airfares.df) + geom_point(aes(x = COUPON, y = FARE), colour = "navy", alpha = 0.7)
ggplot(airfares.df) + geom_point(aes(x = E_INCOME, y = FARE), colour = "navy", alpha = 0.7)
ggplot(airfares.df) + geom_point(aes(x = E_POP, y = FARE), colour = "navy", alpha = 0.7)
```

## explore % of categorical predictors
```{r}
vacation<-prop.table(table(airfares.df$VACATION))
vacation
sw<-prop.table(table(airfares.df$SW))
sw
slot<-prop.table(table(airfares.df$SLOT))
slot
gate<-prop.table(table(airfares.df$GATE))
gate
```

## aggregate and pivot table
```{r}
fare_mean<-aggregate(airfares.df$FARE,by=list(VACATION=airfares.df$VACATION, 
                                    SW=airfares.df$SW,
                                    SLOT=airfares.df$SLOT,
                                    GATECONDITION=airfares.df$GATE), FUN=mean) View(fare_mean)
aggregate(airfares.df$FARE,by=list(VACATION=airfares.df$VACATION),FUN=mean)
aggregate(airfares.df$FARE,by=list(SW=airfares.df$SW),FUN=mean)
aggregate(airfares.df$FARE,by=list(SLOT=airfares.df$SLOT),FUN=mean)
aggregate(airfares.df$FARE,by=list(GATECONDITION=airfares.df$GATE),FUN=mean)


library(reshape) 
mlt.airfares <- melt(airfares.df, id=c("VACATION", "SW","SLOT","GATE"), measure=c("FARE"))
View(mlt.airfares)
cast(mlt.airfares, VACATION ~ SW , subset=variable=="FARE", 
     margins=c("grand_row", "grand_col"), mean)
```

## create dummy variables
```{r}
library(dummies)
airfares.df <- airfares.df[,-c(1,2,3,4)]
airfares.dummy <- dummy.data.frame(airfares.df, sep = ".")
head(airfares.dummy)  
```

## Partition data into traning and validatioin
```{r}
set.seed(121)  
train.rows <- sample(rownames(airfares.dummy), dim(airfares.dummy)[1]*0.6)
train.airfares <- airfares.dummy[train.rows, ]
View(train.airfares)
valid.rows <- setdiff(rownames(airfares.dummy), train.rows) 
valid.airfares <- airfares.dummy[valid.rows, ]
airfares.lm <- lm(FARE ~ ., data = train.airfares)
summary(airfares.lm)
```

## stepwise regression
```{r}
library(stats)
airfares.lm.step <- step(airfares.lm, direction = "backward")
summary(airfares.lm.step)  
airfares.lm.step.pred <- predict(airfares.lm.step, valid.airfares)
accuracy(airfares.lm.step.pred, valid.airfares$FARE)
```

## exhaustive search
```{r}
library(leaps)
airfares.search <- regsubsets(FARE ~ ., data = train.airfares, nbest = 1, nvmax = 10, method = "exhaustive")
sum<-summary(airfares.search)
View(sum)
sum$which
sum$rsq
sum$adjr2
sum$Cp
airfares.lm.search <- lm(FARE ~ VACATION.No+SW.Yes+HI+S_INCOME+E_INCOME+S_POP+E_POP+ SLOT.Controlled+GATE.Free+DISTANCE+PAX, data = train.airfares)
summary(airfares.lm.search)  
airfares.lm.search.pred <- predict(airfares.lm.search, valid.airfares)
accuracy(airfares.lm.search.pred, valid.airfares$FARE)
```

## predict new data
```{r}
newdata<- data.frame(1.202,3,1,0,4442.141,28760,27664,4557004,3195503,0,1,12782,1976, 0)
names(newdata)<-c("COUPON","NEW","VACATION.No","SW.Yes","HI","S_INCOME","E_INCOME","S_POP","E_POP","SLOT.Controlled","GATE.Free","PAX","DISTANCE","FARE")

newprediction <- predict(airfares.lm.search, newdata)
View(newprediction)

newdata1<- data.frame(1.202,3,1,1,4442.141,28760,27664,4557004,3195503,0,1,12782,1976, 0)
names(newdata1)<-c("COUPON","NEW","VACATION.No","SW.Yes","HI","S_INCOME","E_INCOME","S_POP","E_POP","SLOT.Controlled","GATE.Free","PAX","DISTANCE","FARE")

newprediction1 <- predict(airfares.lm.search, newdata1)
View(newprediction1)
```

## select only available factors
```{r}
View(airfares.df)
airfares_new <- airfares.df[ , c("VACATION","S_INCOME","E_INCOME","S_POP","E_POP","DISTANCE","FARE")]

set.seed(121)  
train.rows2 <- sample(rownames(airfares_new), dim(airfares_new)[1]*0.6)
train.airfares2 <- airfares_new[train.rows2, ]
valid.rows2 <- setdiff(rownames(airfares_new), train.rows2) 
valid.airfares2 <- airfares_new[valid.rows2, ]

airfares.search2 <- regsubsets(FARE ~ ., data = train.airfares2, nbest = 1, nvmax = 10, method = "exhaustive")
sum2<-summary(airfares.search2)
View(sum2)
sum2$which
sum2$rsq
sum2$adjr2
sum2$Cp

airfares.lm.search2 <- lm(FARE ~ VACATION+S_INCOME+E_INCOME+S_POP+E_POP+ DISTANCE, data = train.airfares2)
summary(airfares.lm.search2)  
```

## use search2 model to predict new data
```{r}
newdata2<- data.frame("COUPON"=1.202,"NEW"=3,"VACATION"="No","SW"="No","HI"=4442.141,"S_INCOME"=28760,"E_INCOME"=27664,"S_POP"=4557004,"E_POP"=3195503,"SLOT"="Free","GATE"="Free","PAX"=12782,"DISTANCE"=1976, "FARE"=0)
View(newdata2)
newprediction2 <- predict(airfares.lm.search2, newdata2)
View(newprediction2)
```

## Compare accuracy of search model and search2 model
```{r}
airfares.lm.search.pred <- predict(airfares.lm.search, valid.airfares)
accuracy(airfares.lm.search.pred, valid.airfares$FARE)
airfares.lm.search.pred2 <- predict(airfares.lm.search2, valid.airfares2)
accuracy(airfares.lm.search.pred2, valid.airfares2$FARE)
```

